Leadership across cultures and contexts: an exploratory analysis

### 2020-05-17 leadership_across_cultures_contexts.Rmd
Fixed a few spelling errors. Added percentage.

## The benefits and costs of leadership across cultures and contexts
<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership and followership generally assume that, on average, the fitness benefits of leadership outweigh the fitness costs. However, the costs of leadeship and followership have received less attention. Group responsibilities entail opportunity costs. Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. In the absence of institutionalized authority, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile. Physically formidable individuals with strong social networks are likely to be able to bear the costs of leadership more so than their counterparts [@glowacki_leadership_2015].

@@

We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex. Specifically, we fit separate multiple logistic mixed effects regression models of each dimension as a function of these four factors, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts. Models of leadership dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary [@burnham2002].

@@

@@

To test this idea, we operationalized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)). We then calculated the fraction of text records with evidence for one or more of the five elements. See Figure \@ref(fig:fivefold).

@@

Figure \@ref(fig:fivefold) reveals that descriptions of leadership in the ethnographic record includes a heterogeneous mix of these five elements. In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs. Shamans are associated with evidence for *Cognitive capital* and imposing costs.

>>>>

## The benefits and costs of leadership across cultures and contexts
<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership and followership generally assume that, on average, the fitness benefits of leadership outweigh the fitness costs. However, the costs of leadership and followership have received less attention. Group responsibilities entail opportunity costs. Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. In the absence of institutionalized authority, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile. Physically formidable individuals with strong social networks are likely to be able to bear the costs of leadership more so than their counterparts [@glowacki_leadership_2015].

@@

We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex. Specifically, we fit separate multiple logistic mixed effects regression models of each dimension as a function of these four factors, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts. Models whose AIC values were at least two less than the intercept-only models were deemed to vary by those factors [@burnham2002].

@@

```{r capital}
capital_pct <- signif(100*sum(rowSums(df_fivefold2[1:5])>0)/nrow(df_fivefold2), 2)
```

@@

To test this idea, we operationalized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)). About half of the text records (`r capital_pct`%) had evidence for one or more of the five elements, in heterogeneous combinations. See Figure \@ref(fig:fivefold). In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs. Shamans are associated with evidence for *Cognitive capital* and imposing costs.

@@
